# Top 50 ggplot2 Visualizations {#sec-top50ggplot2en}

::: callout-note
This article and code is from [r-statistics.co](https://r-statistics.co/Top50-Ggplot2-Visualizations-MasterList-R-Code.html)
:::

查看简体中文版本: Top 50 ggplot2 可视化 (@sec-top50ggplot2zh)

**An effective chart is one that:**

1.  Conveys the right information without distorting facts.
2.  Is simple but elegant. It should not force you to think much in order to get it.
3.  Aesthetics supports information rather that overshadow it.
4.  Is not overloaded with information

The list below sorts the visualizations based on its primary purpose. Primarily, there are 8 types of objectives you may construct plots. So, before you actually make the plot, try and figure what findings and relationships you would like to convey or examine through the visualization. Chances are it will fall under one (or sometimes more) of these 8 categories.

**These packages are used in this article:**

```{r}
#| label: loading required packages
#| warning: false
library(tidyverse)
library(ggalt)
library(ggExtra)
library(scales)
library(ggthemes)
library(treemapify)
library(ggdendro)
library(ggfortify)
```

## Correlation

The following plots help to examine how well correlated two variables are.

### Scatterplot

The most frequently used plot for data analysis is undoubtedly the scatterplot. Whenever you want to understand the nature of relationship between two variables, invariably the first choice is the scatterplot.

```{r}
#| warning: false
#| label: Scatterplot
theme_set(theme_bw())
options(scipen = 999)
ggplot(midwest, aes(x = area, y = poptotal)) +
  geom_point(aes(color = state, size = popdensity)) +
  geom_smooth(method = "loess", se = FALSE) +
  xlim(c(0, 0.1)) +
  ylim(c(0, 500000)) +
  labs(
    x = "Area",
    y = "Population",
    title = "Scatterplot",
    subtitle = "Area vs Population",
    caption = "Source: midwest"
  )
```

### Scatterplot With Encircling

When presenting the results, sometimes I would encirlce certain special group of points or region in the chart so as to draw the attention to those peculiar cases. This can be conveniently done using the `geom_encircle()` in `ggalt` package.

```{r}
#| label: Scatterplot With Encircling
#| warning: false

midwest_select <- midwest[
  midwest$poptotal > 350000 &
    midwest$poptotal <= 500000 &
    midwest$area > 0.01 &
    midwest$area < 0.1,
]

# Plot
ggplot(midwest, aes(x = area, y = poptotal)) +
  geom_point(aes(col = state, size = popdensity)) + # draw points
  geom_smooth(method = "loess", se = F) +
  xlim(c(0, 0.1)) +
  ylim(c(0, 500000)) + # draw smoothing line
  geom_encircle(aes(x = area, y = poptotal),
    data = midwest_select,
    color = "red",
    size = 2,
    expand = 0.08
  ) + # encircle
  labs(
    subtitle = "Area Vs Population",
    y = "Population",
    x = "Area",
    title = "Scatterplot + Encircle",
    caption = "Source: midwest"
  )
```

### Counts Chart

The second option to overcome the problem of data points overlap is to use what is called a counts chart. Whereever there is more points overlap, the size of the circle gets bigger.

```{r}
#| label: Counts Chart

ggplot(mpg, aes(cty, hwy)) +
  geom_count(aes(colour = "tomato3"), show.legend = FALSE) +
  labs(
    subtitle = "mpg: city vs highway mileage",
    y = "hwy",
    x = "cty",
    title = "Counts Plot"
  )

```

### Bubble plot

While scatterplot lets you compare the relationship between 2 continuous variables, bubble chart serves well if you want to understand relationship within the underlying groups based on:

A Categorical variable (by changing the color) and Another continuous variable (by changing the size of points). In simpler words, bubble charts are more suitable if you have 4-Dimensional data where two of them are numeric (X and Y) and one other categorical (color) and another numeric variable (size).

The bubble chart clearly distinguishes the range of `displ` between the manufacturers and how the slope of lines-of-best-fit varies, providing a better visual comparison between the groups.

```{r}
#| label: Bubble plot
#| warning: false

mpg_select <- mpg[mpg$manufacturer %in% c("audi", "ford", "honda", "hyundai"), ]
ggplot(mpg_select, aes(displ, cty)) +
  geom_jitter(aes(col = manufacturer, size = hwy)) +
  geom_smooth(aes(col = manufacturer), method = "lm", se = FALSE) +
  labs(
    subtitle = "mpg: Displacement vs City Mileage",
    title = "Bubble chart"
  )

```

### Marginal Histogram / Boxplot

If you want to show the relationship as well as the distribution in the same chart, use the marginal histogram. It has a histogram of the X and Y variables at the margins of the scatterplot.

This can be implemented using the `ggMarginal()` function from the `ggExtra` package. Apart from a histogram, you could choose to draw a marginal boxplot or density plot by setting the respective type option.

```{r}
#| label: Marginal Histogram
#| warning: false

# mpg_select <- mpg[mpg$hwy >= 35 & mpg$cty > 27, ]
g <- ggplot(mpg, aes(cty, hwy)) +
  geom_count() +
  geom_smooth(method = "lm", se = FALSE)

ggMarginal(g, type = "histogram", fill = "transparent")
```

## Deviation

Compare variation in values between small number of items (or categories) with respect to a fixed reference.

### Diverging bars

Diverging Bars is a bar chart that can handle both negative and positive values. This can be implemented by a smart tweak with `geom_bar()`. But the usage of `geom_bar()` can be quite confusing. Thats because, it can be used to make a bar chart as well as a histogram. Let me explain.

By default, `geom_bar()` has the stat set to count. That means, when you provide just a continuous X variable (and no Y variable), it tries to make a histogram out of the data.

In order to make a bar chart create bars instead of histogram, you need to do two things.

Set `stat=identity` provide both x and y inside `aes()` where, x is either character or factor and y is numeric. In order to make sure you get diverging bars instead of just bars, make sure, your categorical variable has 2 categories that changes values at a certain threshold of the continuous variable. In below example, the mpg from mtcars dataset is normalised by computing the z score. Those vehicles with mpg above zero are marked green and those below are marked red.

```{r}
#| label: Diverging Bars
#| warning: false
mtcars_new <- mtcars |>
  mutate(
    car_name = factor(rownames(mtcars)),
    mpg_z = round((mpg - mean(mpg)) / sd(mpg), 2), # compute normalized mpg
    mpg_type = ifelse(mpg_z < 0, "below", "above")
  ) |> # above / below avg flag
  arrange(mpg_z) |>
  as_tibble()
# Diverging Barcharts
ggplot(mtcars_new, aes(x = car_name, y = mpg_z, label = mpg_z)) +
  geom_bar(stat = "identity", aes(fill = mpg_type), width = 0.5) +
  scale_fill_manual(
    name = "Mileage",
    labels = c("Above Average", "Below Average"),
    values = c("above" = "#00ba38", "below" = "#f8766d")
  ) +
  labs(
    subtitle = "Normalised mileage from 'mtcars'",
    title = "Diverging Bars"
  ) +
  # using sorted car_name to adjust the rank of the x labels
  scale_x_discrete(limits = mtcars_new$car_name) +
  coord_flip()
```

### Diverging Lollipop Chart

Lollipop chart conveys the same information as bar chart and diverging bar. Except that it looks more modern. Instead of `geom_bar()`, I use `geom_point()` and `geom_segment()` to get the lollipops right. Let’s draw a lollipop using the same data I prepared in the previous example of diverging bars.

```{r}
#| label: Diverging Lollipop Chart
#| warning: false

ggplot(mtcars_new, aes(x = car_name, y = mpg_z, label = mpg_z)) +
  geom_point(stat = "identity", fill = "black", size = 6) +
  # geom_segment draw a line between
  # (x,y) and (xend,yend)
  geom_segment(aes(
    y = 0, x = car_name,
    yend = mpg_z, xend = car_name
  ), color = "black") +
  # label mpg_z -> geom_text
  geom_text(color = "white", size = 2) +
  labs(
    title = "Diverging Lollipop Chart",
    subtitle = "Normalized mileage from 'mtcars': Lollipop"
  ) +
  ylim(-2.5, 2.5) +
  scale_x_discrete(limits = mtcars_new$car_name) +
  coord_flip()

```

### Diverging Dot Plot

Dot plot conveys similar information. The principles are same as what we saw in Diverging bars, except that only point are used. Below example uses the same data prepared in the diverging bars example.

```{r}
#| label: Diverging Dot Plot
#| warning: false

ggplot(mtcars_new, aes(x = car_name, y = mpg_z, label = mpg_z)) +
  geom_point(stat = "identity", aes(color = mpg_type), size = 6) +
  scale_color_manual(
    name = "Mileage",
    labels = c("Above Average", "Below Average"),
    values = c("above" = "#00ba38", "below" = "#f8766d")
  ) +
  geom_text(color = "white", size = 2) +
  labs(
    title = "Diverging Dot Plot",
    subtitle = "Normalized mileage from 'mtcars': Dot plot"
  ) +
  scale_x_discrete(limits = mtcars_new$car_name) +
  coord_flip()

```

## Ranking

Used to compare the position or performance of multiple items with respect to each other. Actual values matters somewhat less than the ranking.

### Dot Plot

Dot plots are very similar to lollipops, but without the line and is flipped to horizontal position. It emphasizes more on the rank ordering of items with respect to actual values and how far apart are the entities with respect to each other.

```{r}
#| label: Dot Plot
#| warning: false

# Prepare data: group mean city mileage by manufacturer.
cty_mpg <- mpg |>
  group_by(manufacturer) |>
  summarise(mileage = mean(cty, na.rm = TRUE)) |>
  rename(make = manufacturer) |>
  mutate(make = factor(make, levels = make)) |>
  arrange(mileage)

ggplot(cty_mpg, aes(x = make, y = mileage)) +
  geom_point(color = "tomato2", size = 3) +
  geom_segment(
    aes(
      x = make, y = min(mileage),
      xend = make, yend = max(mileage)
    ),
    linetype = "dashed", linewidth = 0.1
  ) +
  labs(
    title = "Dot Plot",
    subtitle = "Make Vs Avg. Mileage",
    caption = "source: mpg"
  ) +
  scale_x_discrete(limits = cty_mpg$make) +
  coord_flip()

```

### Slope Chart

Slope charts are an excellent way of comparing the positional placements between 2 points on time. At the moment, there is no builtin function to construct this. Following code serves as a pointer about how you may approach this.

```{r}
#| label: Slope Chart
#| warning: false

# Prepare data
df <- read_csv("data/top50ggplot2/gdppercap.csv")

left_label <- paste(df$continent, round(df$`1952`), sep = ",")
right_label <- paste(df$continent, round(df$`1957`), sep = ",")
df$class <- ifelse(df$`1957` - df$`1952` < 0, "red", "green")

# Plot
ggplot(df) +
  geom_segment(aes(x = 1, y = `1952`, xend = 2, yend = `1957`, color = class),
    linewidth = 0.75, show.legend = FALSE
  ) +
  geom_vline(xintercept = 1, linetype = "dashed", linewidth = 0.1) +
  geom_vline(xintercept = 2, linetype = "dashed", linewidth = 0.1) +
  scale_color_manual(
    labels = c("Up", "Down"),
    values = c("green" = "#00ba38", "red" = "#f8766d")
  ) +
  labs(x = "", y = "Mean GdpPerCap") +
  xlim(0.5, 2.5) +
  ylim(0, (1.1 * max(df$`1957`, df$`1952`))) +
  geom_text(
    label = left_label, y = df$`1952`,
    x = rep(1, nrow(df)), hjust = 1.1, size = 3.5
  ) +
  geom_text(
    label = right_label, y = df$`1957`,
    x = rep(2, nrow(df)), hjust = -0.1, size = 3.5
  ) +
  geom_text(
    label = "Time 1", x = 1,
    y = 1.1 * max(df$`1957`, df$`1952`),
    hjust = 1.2, size = 5
  ) +
  geom_text(
    label = "Time 2", x = 2,
    y = 1.1 * max(df$`1957`, df$`1952`),
    hjust = -0.2, size = 5
  ) +
  theme_classic() +
  theme(
    panel.background = element_blank(),
    panel.grid = element_blank(),
    axis.ticks = element_blank(),
    axis.text.x = element_blank(),
    panel.border = element_blank(),
    plot.margin = unit(c(1, 2, 1, 2), "cm")
  )

```

### Dumbbell Plot

Dumbbell charts are a great tool if you wish to: 1. Visualize relative positions (like growth and decline) between two points in time. 2. Compare distance between two categories.

In order to get the correct ordering of the dumbbells, the Y variable should be a factor and the levels of the factor variable should be in the same order as it should appear in the plot.

```{r}
#| label: Dumbbell Plot
#| warning: false

# Prepare data
health <- read.csv("data/top50ggplot2/health.csv")

health$Area <- factor(health$Area,
  levels = as.character(health$Area)
)

# Plot
ggplot(health, aes(x = pct_2013, xend = pct_2014, y = Area, group = Area)) +
  geom_dumbbell(
    color = "#a3c4dc", size = 0.75,
    colour_x = "#0e668b"
  ) +
  scale_x_continuous(labels = scales::percent) +
  labs(
    x = NULL,
    y = NULL,
    title = "Dumbbell Chart",
    subtitle = "Pct Change: 2013 vs 2014",
    caption = "Source: https://github.com/hrbrmstr/ggalt"
  ) +
  theme_classic() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    plot.background = element_rect(fill = "#f7f7f7"),
    panel.background = element_rect(fill = "#f7f7f7"),
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.major.x = element_line(colour = "grey50"),
    axis.ticks = element_blank(),
    legend.position = "top",
    panel.border = element_blank()
  )

```

## Distribution

When you have lots and lots of data points and want to study where and how the data points are distributed.

### Tufte Boxplot

Tufte box plot, provided by ggthemes package is inspired by the works of Edward Tufte. Tufte’s Box plot is just a box plot made minimal and visually appealing.

```{r}
#| label: Tufte Boxplot
#| warning: false

ggplot(mpg, aes(manufacturer, cty)) +
  geom_tufteboxplot() +
  theme_tufte() + # from ggthemes
  theme(axis.text.x = element_text(angle = 65, vjust = 0.6)) +
  labs(
    title = "Tufte Styled Boxplot",
    subtitle = "City Mileage grouped by Class of vehicle",
    caption = "Source: mpg",
    x = "Class of Vehicle",
    y = "City Mileage"
  )

```

### Population Pyramid

Population pyramids offer a unique way of visualizing how much population or what percentage of population fall under a certain category. The below pyramid is an excellent example of how many users are retained at each stage of a email marketing campaign funnel.

```{r}
#| label: Population Pyramid
#| warning: false

options(scipen = 999)

email_campaign_funnel <- read.csv("data/top50ggplot2/email_campaign_funnel.csv")

# X Axis Breaks and Labels
brks <- seq(-15000000, 15000000, 5000000)
lbls <- paste0(as.character(c(seq(15, 0, -5), seq(5, 15, 5))), "m")

ggplot(
  email_campaign_funnel,
  aes(x = Stage, y = Users, fill = Gender)
) + # Fill column
  geom_bar(stat = "identity", width = 0.6) + # draw the bars
  scale_y_continuous(
    breaks = brks, # Breaks
    labels = lbls
  ) + # Labels
  coord_flip() + # Flip axes
  labs(title = "Email Campaign Funnel") +
  theme_tufte() + # Tufte theme from ggfortify
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.ticks = element_blank()
  ) + # Centre plot title
  scale_fill_brewer(palette = "Dark2") # Color palette

```

## Composition

### Waffle Chart

Waffle charts is a nice way of showing the categorical composition of the total population. Though there is no direct function, it can be articulated by smartly maneuvering the ggplot2 using `geom_tile()` function. The below template should help you create your own waffle.

```{r}
#| label: Waffle Chart
#| warning: false

var <- mpg$class
df <- expand.grid(y = 1:10, x = 1:10)
categ_table <- round(table(var) * ((10 * 10) / (length(var))))
df$category <- factor(rep(names(categ_table), categ_table))

ggplot(df, aes(x = x, y = y, fill = category)) +
  geom_tile(color = "black", linewidth = 0.5) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0), trans = "reverse") +
  scale_fill_brewer(palette = "Set2") +
  labs(
    title = "Waffle Chart", subtitle = "'Class' of vehicles",
    caption = "Source: mpg"
  ) +
  theme(
    plot.title = element_text(size = rel(1.2)),
    axis.text = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank(),
    legend.title = element_blank(),
    legend.position = "right"
  )

```

### Treemap

Treemap is a nice way of displaying hierarchical data by using nested rectangles. The treemapify package provides the necessary functions to convert the data in desired format (treemapify) as well as draw the actual plot (ggplotify).

In order to create a treemap, the data must be converted to desired format using `treemapify()`. The important requirement is, your data must have one variable each that describes the area of the tiles, variable for fill color, variable that has the tile’s label and finally the parent group.

Once the data formatting is done, just call `ggplotify()` on the treemapified data.

```{r}
#| label: Treemap
#| warning: false

proglangs <- read.csv("data/top50ggplot2/proglanguages.csv")

ggplot(proglangs, aes(
  area = value,
  fill = parent, group = parent, label = id
)) +
  geom_treemap() +
  geom_treemap_text(
    fontface = "italic", colour = "white", place = "centre",
    grow = TRUE
  ) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_brewer(palette = "Dark2")

```

## Groups

### Hierarchical Dendrogram

```{r}
#| label: Hierarchical Dendrogram
#| warning: false

theme_set(theme_bw())

hc <- hclust(dist(USArrests), "ave")  # hierarchical clustering

# plot
ggdendrogram(hc, rotate = TRUE, size = 2)

```

### Clusters

It is possible to show the distinct clusters or groups using `geom_encircle()`. If the dataset has multiple weak features, you can compute the principal components and draw a scatterplot using PC1 and PC2 as X and Y axis.

```{r}
#| label: Clusters
#| warning: false

theme_set(theme_classic())

# Compute data with principal components ------------------
df <- iris[c(1, 2, 3, 4)]
pca_mod <- prcomp(df) # compute principal components

# Data frame of principal components ----------------------
df_pc <- data.frame(pca_mod$x,
  Species = iris$Species
) # dataframe of principal components
df_pc_vir <- df_pc[df_pc$Species == "virginica", ] # df for 'virginica'
df_pc_set <- df_pc[df_pc$Species == "setosa", ] # df for 'setosa'
df_pc_ver <- df_pc[df_pc$Species == "versicolor", ] # df for 'versicolor'

# Plot ----------------------------------------------------
ggplot(df_pc, aes(PC1, PC2, col = Species)) +
  geom_point(aes(shape = Species), size = 2) + # draw points
  labs(
    title = "Iris Clustering",
    subtitle = "With principal components PC1 and PC2 as X and Y axis",
    caption = "Source: Iris"
  ) +
  coord_cartesian(
    xlim = 1.2 * c(min(df_pc$PC1), max(df_pc$PC1)),
    ylim = 1.2 * c(min(df_pc$PC2), max(df_pc$PC2))
  ) + # change axis limits
  geom_encircle(data = df_pc_vir, aes(x = PC1, y = PC2)) + # draw circles
  geom_encircle(data = df_pc_set, aes(x = PC1, y = PC2)) +
  geom_encircle(data = df_pc_ver, aes(x = PC1, y = PC2))

```